---
title: "Homework 10"
author: "Grace Smith"
date: "2023-04-05"
output: html_document
---

## Question 1


Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.

```{r}
##############################
# FUNCTION: calculate_zeroes
# packages: none
# purpose:calculate number of zeroes in a numeric vector
# input: 
# output:
# ---------------------------------

calculate_zeroes <- function(x=c(0,5,8,0,7,0,4)) {
  counter <- 0
  for (i in seq_along(x)){
    if(x[i] == 0)
      counter <- counter + 1
  
  }
  return(counter)
}



```

```{r}
calculate_zeroes()
```

## Question 2

Use subsetting instead of a loop to rewrite the function as a single line of code.

```{r}
x <- c(0,5,8,0,7,0,4)

length((x[x == 0]))
```

## Question 3

Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
##############################
# FUNCTION: multi_matrix
# packages: none
# purpose:create matrix in which each element is the product of the row number X column number
# input:
# output:
# ---------------------------------
multi_matrix <- function() {
  m <- matrix(nrow=3, ncol = 3)
  for (i in 1:nrow(m)) {
    for (j in 1:ncol(m)) {
      m[i,j] <- i*j
      
    }
    
  }
print(m)
}
```


```{r}
multi_matrix()
```


## Question 4

In the next few lectures, you will learn how to do a randomization test on your data. We will complete some of the steps today to practice calling custom functions within a for loop. Use the code from the March 31st lecture (Randomization Tests) to complete the following steps:

Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.

```{r}
library(ggplot2)
set.seed(1000)

trt_group <- c(rep("Control", 4),rep("Treatment1",4), rep("Treatment2", 4))
print(trt_group)

z<- c(runif(4) + 1, runif(4) + 10, runif(4) + 25)


df <- data.frame(trt=trt_group,res=z)
print(df)
```



Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.

Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.



```{r}
##############################
# FUNCTION: shuffle_data
# packages: tidyverse
# purpose:
# input:
# output:
# ---------------------------------

library(tidyverse)
shuffle_data <- function(z=df) {
      rep <- rep(NA,100)
      Control <- rep(NA, 100)
      Treatment1 <- rep(NA, 100)
      Treatment2 <- rep(NA, 100)
      
      new_df<-data.frame(rep,Control,Treatment1,Treatment2)

  
  # for (i in 1:100){
  #   new_df[i,1] <- i
  for (i in 1:100){
    new_df[i,1] <- i

    df$res <- sample(df$res)
    
    a<-mean(df[1:4,2])
    new_df[i,2]<- a
    
    b<-mean(df[5:8,2])
    new_df[i,3] <- b
    
    c<-mean(df[9:12,2])
    new_df[i,4] <- c

    }

return(new_df)
  
}
  

shuffle_data(df)



```


Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?


```{r}
new_df <- shuffle_data(df)

qplot(Control, data = new_df, geom = "histogram")

qplot(Treatment1, data = new_df, geom = "histogram")

qplot(Treatment2, data = new_df, geom = "histogram")


```



