---
title: "homework 7"
author: "Grace Smith"
date: "2023-03-08"
output: html_document
---

## Open libraries

```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation


```

## Read in data vector

```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

#z <- rnorm(n=3000,mean=0.2)
#z <- data.frame(1:3000,z)
#names(z) <- list("ID","myVar")
#z <- z[z$myVar>0,]
#str(z)
#summary(z$myVar)


z <- read.table("my_data.csv",header=TRUE,sep=",") #importing dataset
names(z) <- list("ObjectID", "Species", "DBH_2021") #rename columns
z<-subset(z, Species == "ABBA") #only looking Balsam Fir
z<-subset(z, DBH_2021 < 100) # get rid of false, large DBH
z<-subset(z, DBH_2021 > 10.2) # only include trees over overstory size class
View(z)
str(z)
summary(z)
```

## Plot histogram of data

```{r}
p1 <- ggplot(data=z, aes(x=DBH_2021, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)
print(p1)
```

## Add empirical density curve

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```
## Get maximum likelihood parameters for normal


```{r}
normPars <- fitdistr(z$DBH_2021,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute


```

## Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$DBH_2021),len=length(z$DBH_2021))

stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$DBH_2021), args = list(mean = meanML, sd = sdML))
 p1 + stat
```

## Plot exponential probability density
```{r}
expoPars <- fitdistr(z$DBH_2021,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$DBH_2021), args = list(rate=rateML))

p1 + stat + stat2
```

## Plot uniform probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$DBH_2021), args = list(min=min(z$DBH_2021), max=max(z$DBH_2021)))

p1 + stat + stat2 + stat3

```
## Plot gamma probability density
```{r}
gammaPars <- fitdistr(z$DBH_2021,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$DBH_2021), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4

```

## Plot beta probability density

```{r}
pSpecial <- ggplot(data=z, aes(x=DBH_2021/(max(DBH_2021 + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$DBH_2021/max(z$DBH_2021 + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$DBH_2021), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

## Best fit?

Gamma distribution is the best fit



## Simulate new data
```{r}
q<- rgamma(11582, 12.9, rate = 0.775)
q <- data.frame(1:11582,q)
names(q) <- list("ID","myVar")


q1 <- ggplot(data=q, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2)


QgammaPars <- fitdistr(q$myVar,"gamma")
QshapeML <- QgammaPars$estimate["shape"]
QrateML <- QgammaPars$estimate["rate"]


qstat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="magenta", n = length(q$myVar), args = list(shape=QshapeML, rate=QrateML))

q1 + qstat4 # simulated data

p1 + stat4 # my data



```

## **LOOKS GOOD!**